### 2018-05-29 latex/sc18/algorithm.tex
minor changes

\begin{table*}
\centering
\begin{tabular}{|c|cccc|}
\hline
 & \textbf{Computation} & \textbf{Memory - Local MTTKRP} & \textbf{Communication} & \textbf{Memory - Par.~Algorithm}  \\
\hline
Par-NNCP w/ opt.~proc.~grid & $\frac{IR}{P}$ & $\frac{RI^{1/2}}{P^{1/2}}$& $\frac{NRI^{1/N}}{P^{1/N}}$ & $\frac{NRI^{1/N}}{P^{1/N}}$  \\
Par-NNCP w/ gen.~proc.~grid & & & $R\sum_n \frac{I_n}{P_n}$ & $R\sum_n \frac{I_n}{P_n}$  \\
NoDimTree & $\frac{NIR}{P}$ & $\displaystyle \max_n \frac{RI/I_n}{P/P_n}$  &  & \\
\hline
\end{tabular}
\smallskip
\caption{Leading-order per-iteration costs in terms of computation (flops), communication (words moved), and memory (words).  We ignore constants but omit big-Oh notation for clarity.  The first line corresponds to \cref{alg:Par-NNCP-short} with an optimal choice of processor grid and applying the dimension tree optimization locally.  The second line corresponds to the changes in communication and memory for a general processor grid.  The third line corresponds to the changes in computation and memory if the dimension tree optimization is not applied.}
\label{tab:costs}
\end{table*}

>>>>

\begin{table*}
\centering
\begin{tabular}{|c|cccc|}
\hline
 & \textbf{Computation} & \textbf{Memory - Local MTTKRP} & \textbf{Communication} & \textbf{Memory - Par.~Algorithm}  \\
\hline
Par-NNCP w/ opt.~proc.~grid & $\frac{IR}{P}$ & $\frac{RI^{1/2}}{P^{1/2}}$& $\frac{NRI^{1/N}}{P^{1/N}}$ & $\frac{NRI^{1/N}}{P^{1/N}}$  \\
Par-NNCP w/ gen.~proc.~grid & & & $R\sum_n \frac{I_n}{P_n}$ & $R\sum_n \frac{I_n}{P_n}$  \\
Par-NNCP w/o dim.~tree opt. & $\frac{NIR}{P}$ & $\displaystyle \max_n \frac{RI/I_n}{P/P_n}$  &  & \\
\hline
\end{tabular}
\smallskip
\caption{Leading-order per-iteration costs in terms of computation (flops), communication (words moved), and memory (words).  We ignore constants but omit big-Oh notation for clarity.  The first line corresponds to \cref{alg:Par-NNCP-short} with an optimal choice of processor grid and applying the dimension tree optimization locally.  The second line corresponds to the changes in communication and memory for a general processor grid.  The third line corresponds to the changes in computation and memory if the dimension tree optimization is not applied.}
\label{tab:costs}
\end{table*}

### 2018-05-29 latex/sc18/introduction.tex
added figure and more details regarding dimtrees
minor changes

The CP decomposition is a low-rank approximation of a multi-dimensional array, or tensor, which generalizes matrix approximations like the truncated singular value decomposition. As in Figure \ref{fig:cpdecomposition}, CP decomposition approximates the given input matrix as sum of $k$ rank-1 tensors.  CP is often used for finding hidden patterns, or latent factors, within tensor data, particularly when the goal is to interpret the factors, and it is popular within the signal processing, machine learning, and scientific computing communities.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=3in, height=1.5in]{fig/cpdecomposition.pdf}
\caption{CP Decomposition}
\label{fig:cpdecomposition}
\end{center}
\end{figure}

@@

As we detail in the related work, the general techniques for reducing computation and communication are not new.
The recomputation avoidance was proposed in a sequential algorithm \cite{PTC13a}, and the parallelization scheme was proposed and analyzed for general tensors \cite{BKR17-TR} and the algorithm was implemented for 3D tensors \cite{LK+17b}.
The main contribution of this paper is the combination and implementation of these ideas for general $N$D tensors, and we demonstrate superior performance to the existing implementation for 3D tensors \cite{LK+17b}.

We summarize our main contributions as follows:
\begin{itemize}
		\item we present a computation- and communication-efficient algorithm for NNCP with detailed cost analysis,
		\item our algorithm is designed for general $N$-way tensors,
		\item we demonstrate a performance improvement of up to $2.2\times$ over the existing state-of-the-art parallel software on 3D tensors,
		\item we demonstrate efficient parallel scaling of up to $253\times$ on $512$ cores.
\end{itemize}

>>>>

%The CP decomposition is a low-rank approximation of a multi-dimensional array, or tensor, which generalizes matrix approximations like the truncated singular value decomposition. 
%As in Figure \ref{fig:cpdecomposition}, CP decomposition approximates the given input matrix as sum of $k$ rank-1 tensors.  
The CP decomposition is a low-rank approximation of a multi-dimensional array, or tensor, which generalizes matrix approximations like the truncated singular value decomposition.
It approximates the input tensor by a sum of rank-one tensors, which are outer-products of vectors. 
CP is often used for finding hidden patterns, or latent factors, within tensor data, particularly when the goal is to interpret the factors, and it is popular within the signal processing, machine learning, and scientific computing communities.

%\begin{figure}[htbp]
%\begin{center}
%\includegraphics[width=3in, height=1.5in]{fig/cpdecomposition.pdf}
%\caption{CP Decomposition}
%\label{fig:cpdecomposition}
%\end{center}
%\end{figure}

@@

As we detail in the related work, the general techniques for reducing computation and communication have been used in similar contexts.
The recomputation avoidance was proposed in a sequential algorithm \cite{PTC13a}, the parallelization scheme was proposed and analyzed for general tensors \cite{BKR17-TR}, and the algorithm was implemented for 3D tensors \cite{LK+17b}.
%The main contribution of this paper is the combination and implementation of these ideas for general $N$-D tensors, and we demonstrate superior performance to the existing implementation for 3D tensors \cite{LK+17b}.

We summarize our main contributions as follows:
\begin{itemize}
		\item we present the first distributed-memory parallel implementation of NNCP algorithms for arbitrary-dimension dense tensors,
		\item we optimize the use of dimension trees for dense tensors, avoiding recomputation across multiple MTTKRPs,
		\item our parallel algorithm is communication optimal with a carefully chosen processor grid,
		\item we demonstrate a performance improvement of up to $2.2\times$ over the existing state-of-the-art parallel software on 3D tensors,
		\item our implementation obtains efficient parallel scaling of up to $253\times$ on $512$ cores.
\end{itemize}
